### 2016-11-21 local-capability-machine.tex
removed figures (for now)

@@

\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. This call assumes $r_{\var{stk}}$ contains a stack capability. An overview of this call:
  \begin{itemize}
  \item Push restore code to the stack.
  \item Push return address capability 
  \item Push ``private'' registers to the stack.
  \item Push stack pointer
  \item Create local enter capability for restoration
  \item Restrict stack capability
  \item Clear unused registers
  \item Jump
  \item Upon return: Run restore code
  \item Return address:  Restore ``private'' state
  \end{itemize}
A more detailed description of the above steps:
\begin{description}
  \item [Push restore code to the stack]
    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
  \item [Push ``private'' registers to the stack]
    Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
  \item [Push return address capability]
    Push a capability for the return address (on the heap) to the stack.
  \item [Push stack pointer]
    Push the full stackpointer to the heap.
  \item [Create local enter capability for restoration]
    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
  \item [Restrict stack capability]
    Make the stack capability only govern the unused part.
  \item [Clear unused registers]
    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
  \item [Jump] Jump to $f$.
\lau{same issue as with \texttt{call}, where dos $f$ come from.}
  \item [Restore code]
    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
  \item [Return address: Restore ``private'' state]
    \begin{itemize}
    \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
    \item Pop the restore code of the stack
    \end{itemize}
\end{description}

This call assumes the presence of a stack! Push the code that restores the stack pointer, $\pcreg$ and $r_0$. Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?} . Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$. Restrict the stack pointer to the ``unused part''. Push the arguments to the stack and jump to \texttt{f}.
  \begin{description}
  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code. \dominique{restores other registers too, right?}
  \end{description}
Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
\end{description}

\begin{figure}
  \label{fig:stack-before-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local state\\
   & $\vdots$
\end{tabular}
\caption{Stack after everything stored}
\end{figure}

@@

\begin{figure}
  \label{fig:stack-after-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local state\\
   & $\vdots$
\end{tabular}
\caption{Stack after everything has been restored}
\end{figure}
\subsection{Labels}
\texttt{l:} is a meta level label that can be used to refer to a specific address.

>>>>

\usepackage{rotating}

@@

\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. This call assumes $r_{\var{stk}}$ contains a stack capability. An overview of this call:
  \begin{itemize}
  \item Push restore code to the stack.
  \item Push return address capability 
  \item Push ``private'' registers to the stack.
  \item Push stack pointer
  \item Create local enter capability for restoration
  \item Restrict stack capability
  \item Clear unused registers
  \item Jump
  \item Upon return: Run restore code
  \item Return address:  Restore ``private'' state
  \end{itemize}
A more detailed description of the above steps:
\begin{description}
  \item [Push restore code to the stack]
    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
  \item [Push ``private'' registers to the stack]
    Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
  \item [Push return address capability]
    Push a capability for the return address (on the heap) to the stack.
  \item [Push stack pointer]
    Push the full stackpointer to the heap.
  \item [Create local enter capability for restoration]
    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
  \item [Restrict stack capability]
    Make the stack capability only govern the unused part.
  \item [Clear unused registers]
    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
  \item [Jump] Jump to $f$.
\lau{same issue as with \texttt{call}, where dos $f$ come from.}
  \item [Restore code]
    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
  \item [Return address: Restore ``private'' state]
    \begin{itemize}
    \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
    \item Pop the restore code of the stack
    \end{itemize}
\end{description}
\end{description}
\begin{comment}
\lau{todo: adjust figures}
\begin{figure}
  \label{fig:stack-before-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local stack\\
   & $\vdots$
\end{tabular}
\hspace{0.5cm}
\begin{tabular}{*{10}{>{\centering\arraybackslash}p{0.2cm}}}
  \begin{sideways}
    $\pcreg$
  \end{sideways}
& 
  \begin{sideways}
    $r_{\var{stk}}$
  \end{sideways}
& 
  \begin{sideways}
    $r_0$
  \end{sideways}
& 
  \begin{sideways}
    $r_{\var{args},1}$
  \end{sideways}
&

& 
  \begin{sideways}
    $r_{\var{args},n}$ 
  \end{sideways}
& 
  \begin{sideways}
    $r_{\var{priv},1}$
  \end{sideways}
&

&
  \begin{sideways}
    $r_{\var{priv},n}$
  \end{sideways}
&\\
\hline
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$\pc$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$c_s$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$c_0$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$w_1$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
 $\phantom{1}$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$w_n$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$v_1$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
  $\phantom{1}$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
$v_n$
} &
\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm} | }{
  $\phantom{1}$
}\\
\hline
\end{tabular}
\caption{Stack after everything stored}
\end{figure}

@@

\begin{figure}
  \label{fig:stack-after-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local state\\
   & $\vdots$
\end{tabular}
\caption{Stack after everything has been restored}
\end{figure}
\end{comment}
\subsection{Labels}
\texttt{l:} is a meta level label that can be used to refer to a specific address.
